The expected shape of the Milky Way's Dark Matter halo

### 2019-02-13 Paper/paper.tex
edit

There are two main ways to estimate the DM halo shape at a fixed
radius: by computing isopotential or isodensity surfaces.
Observational inference models estimate the shape from the
isopotential contours, while simulations work with the isodensity
contours which can be  directly calculated from particle positions.  
However, the density contours are not smooth and are very 
sensitive to the presence of small satelites.
For this reason we choose to measure the shape by taking
volume-enclosed particles, rather than shell-enclosed.

We follow the shape measurement method presented by
\cite{Allgood_et_al._2006} that uses the reduced
inertia tensor,

@@

We start the calculations taking into account particles within a
sphere of radius $R$ and then  recharacterize the triaxial parameters by taking into account
particles within an ellipsoid of semi-axes $r,r/q,r/s$ and weighted
distance $d^2=x^2+(y/q)^2+(z/s)^2$, where $q = b/a$ and $s=c/a$
are the previously calculated axial ratios. 
We repeat this process until the average deviation of semi-axes is
less than $10^{-6}$.  
This is the same method used to estimate the halo shape in the DM-only
Aquarius simulations \citep{Vera-Ciro_et_al._2011}.

@@

\subsection{Radial trends}

In DM-only simulations halos are monotonically rounder with increasing
radius, this confirms results already reported in the literature
\citep{Vera-Ciro_et_al._2011}. 
Figure \ref{fig:slices} illustrates this effect.
There we show the DM density of a DM-only halo at redshift zero.
The halo is aligned with the minor axis. 
The ellipsoid shows the outermost boundary of the estimated shape ellipsoid.
Formally, this effect is reflected in the axial ratios $b/a,c/a$, which tend to unity
on the outerskirts of the halo. As a representative sample of this
relations, we show on figure \ref{fig:DM relation} the evolution, for
a specific halo, of the axial ratios $b/a,c/a$ and the triaxiality
parameter $T=\frac{a^2-b^2}{a^2-c^2}$ in a wide range of radii.

@@

@@

\begin{figure}
\centering
{\includegraphics[width=0.8\columnwidth]{./pics/halo16.png}}
\caption{Radial profile for axial ratios and the triaxiality parameter $T=\frac{1-b/a}{1-c/a}$ for halo 16.
Radial profile was sampled at redshift 0 and there are two sets of
lines showing MHD and DM-only versions of the axial ratios profile}
\label{fig:DM_MHD}
\end{figure}

>>>>

The DM halo shape at a fixed radius is an estimate of either
the isopotential or isodensity surfaces.  
Observational inference models usually estimate the 
isopotential contours which are probed by tracers (gas, stars), while
simulations work with the isodensity contours which can be directly
calculated from particle positions.

However, in numerical simulations the density contours are not smooth
and are sensitive to the presence of small satelites. 
For this reason we choose to measure the shape by taking
volume-enclosed particles, rather than shell-enclosed.  
We follow the shape measurement method presented by
\cite{Allgood_et_al._2006} that uses the reduced
inertia tensor,

@@

We start the calculations taking into account particles within a
sphere of radius $R$ and then  recharacterize the triaxial parameters
by taking into account particles within an ellipsoid of semi-axes
$r,r/q,r/s$ and re-scaled distance $d^2=x^2+(y/q)^2+(z/s)^2$, where $q
= b/a$ and $s=c/a$ are the previously calculated axial ratios. 
We repeat this process until the average deviation of semi-axes is
less than $10^{-6}$.  
This is the same method used to estimate the halo shape in the DM-only
Aquarius simulations \citep{Vera-Ciro_et_al._2011}.

@@

\subsection{Radial trends at $z=0$}

@@

We find that in DM-only simulations halos are monotonically rounder
with increasing radius, confirming results already reported in the
literature \citep{Vera-Ciro_et_al._2011}. 
Figure \ref{fig:slices} illustrates this effect.
There we show the density a DM-only halo at redshift zero in a thin
slice passing through the halo's centre, each panel shows the halo at
different radii.
The ellipse plotted over the contours indicates the outermost
boundary of the estimated shape ellipsoid. 
Indeed, the ellipse panel on the right (outer halo) is rounder than
the ellipse on the left (inner halo).

We quantify this effect by plotting the axial ratios $b/a$, $c/a$,
$c/b$ and the triaxiality $T\equiv \frac{1-b/a}{1-c/a}$ as a
function of radius. 
As a representative sample of this
relations, we show on figure \ref{fig:triaxial_radius} the evolution, for
a specific halo, of the axial ratios $b/a,c/a$ and the triaxiality
parameter $T=\frac{a^2-b^2}{a^2-c^2}$ in a wide range of radii.

\begin{figure}
\centering
{\includegraphics[width=0.8\columnwidth]{./pics/halo16.png}}
\caption{Radial profile for the axial ratios and the triaxiality parameter
  $T=\frac{1-b/a}{1-c/a}$ at redshift $z=0$. 
  Radial profile was sampled at
  redshift 0 and there are two sets of lines showing MHD and DM-only
  versions of the axial ratios profile} \label{fig:triaxial_radius}
\end{figure}

@@
